---
title: "Exploratory Data Project Pt 1"
author: "Maggie McGlenn"
date: "Date"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=6)
# feel free to change fig.height or fig.width here or in specific chunk headings. 
```


```{r}
# load packages - add any you would like
library(tidyverse)
library(lme4)
library(lmerTest)
library(gridExtra)
library(knitr)
library(corrplot)
library(ellipse)
```

## Abstract 

A 150 word abstract.


\newpage

### Background and Significance   

There are many factors which effect plant growth. 

### Data  

Within a climate zone there are 5 trees and within each tree there are three strata and in each strata one branch was randomly selected and on each branch 5 leafs were randomly selected.
variables relating to leaf are level 1 (Length, Width, Area, petiole width, petiole length, herbivory), variables relating to branch are level 2 (strata, branch length, branch), and variables relating to the whole tree are level 3 (climate zone, tree height, tree girth, species).

Type data section here. Include captions with tables and graphs. Examples:  

```{r}
leaf <- read.csv("https://raw.githubusercontent.com/proback/BeyondMLR/master/data/mudamalai.csv")
leaf<- leaf%>% mutate(Species=factor(Species),
                      Zone=factor(Zone),
                      Strata= factor(Strata),
                      Branch= factor(Branch))
leaf<- leaf[,-c(1)]
T1 <- head(leaf %>% select(where(is.numeric)),10)       
kable(T1, caption="First 10 rows of numeric variables in dataset")
```

```{r, fig.cap= "Correlation between numeric variables" }
Corr <- cor(select_if(leaf, is.numeric), use="complete.obs") 
corrplot.mixed(Corr)
```

 
```{r, fig.cap= "Distribution of the length, width, and area of leaves"}
lengthdist<- ggplot(data=leaf,aes(x=Length)) + 
  geom_histogram( fill = "white",
                 color = "black") + 
  theme_bw() + 
  xlab("Length") + ylab("Frequency") + labs(title="(a)")

WidthDist <- ggplot(data=leaf,aes(x=Width)) + 
  geom_histogram( fill = "white", 
                 color = "black") + 
  theme_bw() + 
  xlab("Width") + ylab("") + labs(title="(b)")
AreaDist <- ggplot(data=leaf,aes(x=Area)) + 
  geom_histogram(fill = "white", 
                 color = "black ") + 
  theme_bw() + 
  xlab("Area") + ylab("") + labs(title="(c)")
grid.arrange(lengthdist, WidthDist, AreaDist,ncol=3)
```

```{r, fig.cap="Relationship between tree height and tree girth by species of tree"}
ggplot(leaf, aes(x=Tree.girth, y=Tree.height, color=Species))+geom_point()+
  geom_smooth(method="lm", se=FALSE)+
  theme_bw()+
  xlab("Tree Girth")+
  ylab("Tree Height")
```



```{r, fig.cap="Relation between the strata where a branch is located and branch length"}
ggplot(leaf, aes(x=Branch.length, y=factor(Strata)))+
  geom_boxplot()+
  theme_bw()+
  ylab("Strata")+
  xlab("Branch Length") 
```



```{r, fig.cap= "Distribution of leaf area for each species by climate zone"}
ggplot(leaf, aes(x=Species, y=Area, color=Zone))+
  geom_boxplot()+
   ylab("Leaf Area")+
  xlab("Species")+
  theme_bw()
```

### Methods  

Type methods section here. Write your final model in equation form, like this:


\begin{align*}
Y_{ij} & = [\alpha_{0}+\alpha_{1}\textrm{Orch}_{i}+\beta_{0}\textrm{LargeEns}_{ij}+\beta_{1}\textrm{Orch}_{i}\textrm{LargeEns}_{ij}] \\
    & + [u_{i}+v_{i}\textrm{LargeEns}_{ij}+\epsilon_{ij}]
\end{align*}

where,

$$
\left[ \begin{array}{c}
            u_{i} \\ v_{i}
          \end{array}  \right] \sim N \left( \left[
          \begin{array}{c}
            0 \\ 0
          \end{array} \right], \left[
          \begin{array}{cc}
            \sigma_{u}^{2} & \rho_{uv}\sigma_{u}\sigma_v \\
            \rho_{uv}\sigma_{u}\sigma_v & \sigma_{v}^{2}
          \end{array} \right] \right) 
$$

and $\epsilon_{ij}\sim N(0,\sigma^2)$.   



### Results  

Write your results section here. Include the fixed and random effects tables, as below. 

```{r}
music <- music %>% mutate(orch = as.numeric(instrument=="orchestral instrument"),
                          large = as.numeric(perform_type=="Large Ensemble"))
M <- lmer(na ~ orch + large + orch:large +
  (large|id), data = music)

m2<- lmer(Area~ Zone+Species)

```

```{r, fig.cap="Variance and Covariance Estimates for Random Effects"}
kable(as.data.frame(VarCorr(M)))
```


```{r, fig.cap="Estimates of Fixed Effects"}
kable(summary(M)$coefficients)
```


### Discussion and Conclusions   


Type discussion section. 

\newpage 

### References  (5 points)

References in APA format. For example:

Sadler, M. E., & Miller, C. J. (2010). Performance anxiety: A longitudinal study of the roles of personality and experience in musicians. Social Psychological and Personality Science, 1(3), 280-287.

Lin, M. C. (2019). An Investigation Of Music Performance Anxiety In Taiwanese Pianists, Vocalists, String And Wind Instrumentalists At The College Level.

Stoeber, J., & Eismann, U. (2007). Perfectionism in young musicians: Relations with motivation, effort, achievement, and distress. Personality and Individual Differences, 43(8), 2182-2192.

Roback, P., & Legler, J. (2021). Beyond Multiple Linear Regression: Applied Generalized Linear Models And Multilevel Models in R. CRC Press.



\newpage

## Appendix (optional)
